---
title: "STAT462 Assignment 2"
author: 
  - "David Ewing (82171165)"
  - "Xia Yu (62380486)"
date: "`r Sys.Date()`"

output:
  pdf_document:
    latex_engine: xelatex
    fig_caption: true
    number_sections: true
    toc: true
    keep_tex: true
    fig_crop: false
    includes:
      in_header: ../doc/fonts.tex

mainfont: Arial          # <-- THIS is inside the YAML, but not under pdf_document
fontsize: 11pt           # <-- Same here
geometry: margin=1in     # <-- And here
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages, include=FALSE}
required_packages <- c(
  "caret", "class", "cowplot", "dplyr", "flextable", "GGally", "ggplot2", 
  "glmnet", "glue", "kableExtra", "kknn", "knitr", "MASS", "officer", 
  "patchwork", "pROC", "rsample", "skimr", "tibble", "tidyverse", "tree", 
  "yardstick"
)

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
  library(pkg, character.only = TRUE)
}

library(conflicted)
if ("dplyr" %in% loadedNamespaces()) {
  try(conflict_prefer("filter", "dplyr"), silent = TRUE)
  try(conflict_prefer("select", "dplyr"), silent = TRUE)
}
```

```{r deebug, include=FALSE}
# ---- All pretty_* helpers from A & B ----

pretty_df <- function(df, title = NULL, fontsize = 10, n = 5) {
  if (!is.data.frame(df)) stop("Input must be a data frame")
  if (n > 0) df <- head(df, n)
  if (n < 0) df <- tail(df, abs(n))

  ft <- flextable::flextable(df) |>
    flextable::fontsize(size = fontsize, part = "all") |>
    flextable::align(align = "center", part = "all") |>
    flextable::theme_booktabs() |>
    flextable::bold(i = 1, part = "header") |>
    flextable::autofit()

  current_width <- sum(ft$body$colwidths)

  if (!is.null(title)) {
    estimated_char_width <- 0.07
    title_width <- nchar(title) * estimated_char_width
    if (title_width > current_width) {
      scale_factor <- title_width / current_width
      new_widths <- ft$body$colwidths * scale_factor
      for (j in seq_along(new_widths)) {
        ft <- flextable::width(ft, j = j, width = new_widths[j])
      }
    }
    ft <- flextable::set_caption(ft, caption = title)
  }

  return(ft)
}

pretty_qda <- function(qda_model) {
  response_var <- as.character(attr(qda_model$terms, "variables")[[2]])
  predictor_vars <- attr(qda_model$terms, "term.labels")
  qda_equation <- paste0("P(", response_var, " = k | ", paste(predictor_vars, collapse = ", "), 
                         ") ∝ π_k × f_k(", paste(predictor_vars, collapse = ", "), ")")

  centroids_df <- as.data.frame(qda_model$means)
  centroids_df[[response_var]] <- rownames(centroids_df)
  centroids_df <- centroids_df[, c(response_var, setdiff(names(centroids_df), response_var))]

  priors_df <- as.data.frame(qda_model$prior)
  priors_df[[response_var]] <- rownames(priors_df)
  colnames(priors_df) <- c("prior_probability", response_var)
  priors_df <- priors_df[, c(response_var, "prior_probability")]

  centro_ft = pretty_df(centroids_df, qda_equation)
  priors_ft = pretty_df(priors_df, qda_equation)

  print(centro_ft)
  print(priors_ft)
  print(qda_equation)

  return(list(
    centroids = centro_ft,
    priors = priors_ft,
    equation = qda_equation
  ))
}

pretty_summary <- \(df) {
  ft <- select(df, where(is.numeric)) |>
    summarise(across(
      everything(),
      .fns = list(
        Mean   = \(x) mean(x, na.rm = TRUE),
        Median = \(x) median(x, na.rm = TRUE),
        Min    = \(x) min(x, na.rm = TRUE),
        Max    = \(x) max(x, na.rm = TRUE),
        IQR    = \(x) IQR(x, na.rm = TRUE),
        nNA    = \(x) sum(is.na(x))
      )
    )) |>
    pivot_longer(cols = everything(), names_to = c("Variable", ".value"), names_sep = "_") |>
    mutate(across(where(is.numeric), round, 2)) |>
    flextable() |>
    set_header_labels(
      Variable = "Feature", Mean = "Mean", Median = "Median",
      Min = "Min", Max = "Max", IQR = "Interquartile Range", nNA = "Missing Values"
    ) |>
    autofit() |>
    theme_vanilla()
}

pretty_ggplot <- \(plot, title = "ggplot Summary") {
  if (!inherits(plot, "gg")) stop("Input must be a ggplot object.")
  plot_data <- tryCatch(plot$data, error = function(e) NULL)
  geoms <- sapply(plot$layers, function(layer) class(layer$geom)[1])
  mappings <- plot$mapping
  global_aes  <- names(mappings)
  global_vals <- sapply(mappings, function(x) rlang::expr_text(x))

  p_title <- plot$labels$title %||% title %||% ""
  x_lab <- plot$labels$x %||% ""
  y_lab <- plot$labels$y %||% ""
  colour_lab <- plot$labels$colour %||% plot$labels$color %||% ""

  df <- data.frame(
    Component = c("Title", "X Axis", "Y Axis", "Colour Legend", "Geoms", global_aes),
    Value     = c(p_title, x_lab, y_lab, colour_lab, paste(geoms, collapse = ","), global_vals),
    stringsAsFactors = FALSE
  )

  ft <- pretty_df(df)
  return(ft)
}
```

# 1. Introduction

This report presents a comprehensive statistical modeling analysis conducted for Assignment 2 of STAT462. The objective is to develop and evaluate predictive models for a multi-class classification problem using color data. The response variable of interest is `colour`, which consists of five categories: red, blue, pink, purple, and brown. The predictor variables are RGB color values (`r` and `b`), extracted from images.

Two modeling approaches are explored and compared:

1. **Quadratic Discriminant Analysis (QDA)** – a generative classification model that fits a separate multivariate normal distribution for each class and allows each class to have its own covariance matrix.
2. **k-Nearest Neighbors (k-NN)** – a non-parametric method that classifies a point based on the majority class among its k nearest neighbors in the feature space.

Throughout the report, these models are evaluated based on their classification accuracy, decision boundaries, and ROC curves (using one-vs-rest strategy) to assess their ability to distinguish each class from the others.

The analysis proceeds by first introducing the dataset, followed by model development, performance evaluation, and interpretation of results.

# 2. Data

```{r data-class-distribution-B, include=TRUE}
# Generate class summary from the training data
print(c_train)
c_class_summary <- c_train %>%
  count(colour, name = "Count") %>%
  mutate(Proportion = round(Count / sum(Count), 3))

# Display formatted table
pretty_df(c_class_summary, title = "Class Distribution in Training Set")
```
